---
title: "Weather and Corn Yield Regressions"
author: "MJ Carlson"
date: "2/25/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(R.matlab)
library(rnassqs)
```

## Weather Data Analysis

### Load the PRISM daily maximum temperatures

```{r tmax data}

# daily max temperature
# dimensions: counties x days x years
prism <- readMat("data/prismiowa.mat")

# look at county #1
t_1981_c1 <- prism$tmaxdaily.iowa[1,,1]
t_1981_c1[366]
plot(1:366, t_1981_c1, type = "l")

ggplot() +
  geom_line(mapping = aes(x=1:366, y = t_1981_c1)) +
  theme_bw() +
  xlab("day of year") +
  ylab("daily maximum temperature (°C)") +
  ggtitle("Daily Maximum Temperature, Iowa County #1")


```


```{r tidying up}

# assign dimension names to tmax matrix
dimnames(prism$tmaxdaily.iowa) <- list(prism$COUNTYFP, 1:366, prism$years)

# converted 3d matrix into a data frame
tmaxdf <- as.data.frame.table(prism$tmaxdaily.iowa)

# relabel the columns
colnames(tmaxdf) <- c("countyfp","doy","year","tmax")
tmaxdf <- tibble(tmaxdf)

```

## Temperature trends

### Summer temperature trends: Winneshiek County

```{r temp trends}

tmaxdf$doy <- as.numeric(tmaxdf$doy)
tmaxdf$year <- as.numeric(as.character(tmaxdf$year))

winnesummer <- tmaxdf %>%
  filter(countyfp==191 & doy >= 152 & doy <= 243) %>%
  group_by(year) %>%
  summarize(meantmax = mean(tmax))

ggplot(winnesummer, mapping = aes(x = year, y = meantmax)) +
  geom_point() +
  theme_bw() +
  labs(x = "year", y = "Tmax (°C)") +
  geom_smooth(method = lm)

lm_summertmax <- lm(meantmax ~ year, winnesummer)
summary(lm_summertmax)

```

### Winter Temperatures - Winneshiek County

```{r winter temps}

winnewinter <- tmaxdf %>%
  filter(countyfp==191 & (doy <= 59 | doy >= 335) & !is.na(tmax)) %>%
  group_by(year) %>%
  summarize(meantmax = mean(tmax))

ggplot(winnewinter, mapping = aes(x = year, y = meantmax)) +
  geom_point() +
  theme_bw() +
  labs(x = "year", y = "Tmax (°C)") +
  geom_smooth(method = lm)

lm_wintertmax <- lm(meantmax ~ year, winnewinter)
summary(lm_wintertmax)

```

### Multiple regression -- Quadratic time trend

```{r quadratic temp trend}

winnewinter$yearsq <- winnewinter$year^2

lm_wintertmaxquad <- lm(meantmax ~ year + yearsq, winnewinter)
summary(lm_wintertmaxquad)
winnewinter$fitted <- lm_wintertmaxquad$fitted.values

ggplot(winnewinter) +
  geom_point(mapping = aes(x = year, y = meantmax)) +
  geom_line(mapping = aes(x = year, y = fitted)) +
  theme_bw() +
  labs(x = "year", y = "tmax")

```

### Download NASS corn yield data

```{r yield download}

# set our API key with NASS
nassqs_auth(key = "DD0322AC-2B3C-34A0-B93C-52D5A8844518")

# parameters to query on 
params <- list(commodity_desc = "CORN", util_practice_desc = "GRAIN", prodn_practice_desc = "ALL PRODUCTION PRACTICES", year__GE = 1981, state_alpha = "IA")

# download
cornyieldsall <- nassqs_yields(params)

cornyieldsall$county_ansi <- as.numeric(cornyieldsall$county_ansi)
cornyieldsall$yield <- as.numeric(cornyieldsall$Value)

# clean and filter this dataset
cornyields <- select(cornyieldsall, county_ansi, county_name, yield, year) %>%
  filter(!is.na(county_ansi) & !is.na(yield))
cornyields <- tibble(cornyields)

```

## Assignment

### Question 1a: Extract Winneshiek County corn yields, fit a linear time trend, make a plot. Is there a significant time trend?

```{r}
# county data
cornWinneshiek<-cornyields%>%
  filter(county_name == "WINNESHIEK")

# linear model
lm_Winneshiek <- lm(yield ~ year, cornWinneshiek)
summary(lm_Winneshiek)

# plot
ggplot(cornWinneshiek, mapping = aes(x = year, y = yield)) +
  geom_point() +
  geom_smooth(method = lm) +
  labs(x = "year", y = "yield")

```


### Question 1b: Fit a quadratic time trend (i.e., year + year^2) and make a plot. Is there evidence for slowing yield growth? 

```{r}
#year^2
cornWinneshiek$yearsq <- cornWinneshiek$year^2

# linear model
lm_Winneshiek <- lm(yield ~ year + yearsq, cornWinneshiek)
summary(lm_Winneshiek)
cornWinneshiek$fitted <- lm_Winneshiek$fitted.values


# plot
ggplot(cornWinneshiek) +
  geom_point(mapping = aes(x = year, y = yield)) +
  geom_line(mapping = aes(x = year, y = fitted)) +
  theme_bw() +
  labs(x = "year", y = "yield")
```


### Question 2 -- Time Series: Let's analyze the relationship between temperature and yields for the Winneshiek County time series. Use data on yield and summer avg Tmax. Is adding year or Tmax^2 to your model helpful? Make a plot and interpret the results.

```{r}

```


### Question 3 -- Cross-Section: Analyze the relationship between temperature and yield across all counties in 2018. Is there a relationship? Interpret the results.

```{r}

```


### Question 4 -- Panel: One way to leverage multiple time series is to group all data into what is called a "panel" regression. Convert the county ID code ("countyfp" or "county_ansi") into factor using as.factor, then include this variable in a regression using all counties' yield and summer temperature data. How does the significance of your temperature coefficients (Tmax, Tmax^2) change? Make a plot comparing actual and fitted yields and interpret the results of your model.

```{r}

```


### Question 5 -- Soybeans: Download NASS data on soybean yields and explore either a time series relationship for a given county, the cross-sectional relationship for a given year, or a panel across all counties and years.

```{r}

```


### Bonus: Find a package to make a county map of Iowa displaying some sort of information about yields or weather. Interpret your map.

### Bonus #2: Challenge question - map trends in corn yields by county across Iowa. Interpret your map.
